The Error Estimate of Force Calculation in the Inhomogeneous Molecular Systems

### 2011-12-15 shortRange.inhomo/alltext/pubilsh/main.tex
add publish

\begin{abstract}
  \recheck{
    In the present paper, we develop an accurate error estimate of the
    non-bonded short-range interaction for the inhomogeneous molecular
    systems.
    The root mean square force error is proved to be decomposed into
    three additive parts: the homogeneity error, the inhomogeneity error
    and the correlation error.
    % The homogeneity error is dominant
    % in the bulk regions.
    The magnitude of the inhomogeneity error, which
    is dominant in the interfacial regions, 
    can be more than one order of magnitude larger than the homogeneity
    error.
    % This is the reason why a direct simulation may waste
    % a huge amount of computational effort.
    This is the reason why a standard simulation with fixed
    cut-off radius is either less accurate if the cut-off is too small,
    or wastes considerable computational effort if the cut-off is too large.
    Therefore, based on the error estimate, the adaptive cut-off
    and long-range force correction methods are proposed to
    boost the efficiency and accuracy of the simulation, respectively.
    The
    effectiveness of the proposed methods are demonstrated by
    molecular dynamics 
    simulations of the liquid-vapor equilibrium and the nanoscale particle
    collision. Different roles of the homogeneity error and inhomogeneity
    error are also discussed.
  }
\end{abstract}

@@

The naive calculation of the convolutions in the error estimate \eqref{eqn:tmp5} requires
a computational cost of $\mathcal O(N^2)$. To estimate the force error
on the fly, the computational cost of the estimate should be reduced to
at least $\mathcal O(N\log N)$.  This is achieved by  using the fast Fourier transform. We
assume the simulation box is uniformly divided in to small bins.  The
positions of these bins are
\begin{align}
  \v r_{i_1, i_2, i_3} & =
  \frac{i_1}{M_1}\v a_1 + 
  \frac{i_2}{M_2}\v a_2 + 
  \frac{i_3}{M_3}\v a_3,\qquad 0\leq i_\alpha< M_\alpha
\end{align}
where $\v a_\alpha$ are box vectors, and $M_\alpha$ is the number of
divisions on each direction. In the reciprocal space, a corresponding
lattice is set up:
\begin{align}
  \v k_{m_1, m_2, m_3} & =
  m_1\v a_1^\ast +
  m_2\v a_2^\ast +
  m_3\v a_3^\ast, \qquad 0\leq m_\alpha<M_\alpha
\end{align}
where $\v a_\alpha^\ast$ are reciprocal box vectors defined by $\v
a_\alpha\cdot\v a_\beta^\ast = \delta_{\alpha\beta}$.  In each bin the
error and the particles number density are assumed to be constant,
then, a Fourier expansion can be obtained:
\begin{align}\nonumber
  \mathcal E^2_{\textrm{homo}}(\v r_{i_1,i_2,i_3})
  & =
  \frac1V\sum_{m_1=-\infty}^\infty\sum_{m_2=-\infty}^\infty\sum_{m_3=-\infty}^\infty
  [{\mathcal E}^2_{\textrm{homo}}]^{\wedge}(\v k_{m_1,m_2,m_3})
  % e^{2\pi i\v k_{m_1,m_2,m_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{m_1i_1}{M_1} + \frac{m_2i_2}{M_2} + \frac{m_3i_3}{M_3}
  \bigg)
  \bigg] \\\label{eqn:tmp13}
  & =
  \frac1V\sum_{m_1=0}^{M_1}\sum_{m_2=0}^{M_2}\sum_{m_3=0}^{M_3}
  \hat{K}_{\textrm{homo}}(\v k'_{m_1,m_2,m_3})\:
  \hat{\rho}(\v k_{m_1,m_2,m_3})
  % e^{2\pi i\v k_{m_1,m_2,m_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{m_1i_1}{M_1} + \frac{m_2i_2}{M_2} + \frac{m_3i_3}{M_3}
  \bigg)
  \bigg].
\end{align}
Here $K_{\textrm{homo}}(\v r)$ denotes $\vert\v f^c(\v r)\vert^2$.
The wedge notation $\wedge$ means the Fourier transform. The prime on $\v k_{m_1,m_2,m_3}$ means
that the periodic image $m_\alpha - M_\alpha$ should use instead of
$m_\alpha$ when $m_\alpha \geq M_\alpha/2$.
The second equation in \eqref{eqn:tmp13}
holds due to the fact:
$[\mathcal E^2]^\wedge = [K_{\textrm{homo}}\ast\rho]^\wedge = \hat{K}_{\textrm{homo}} \times \hat\rho$. The Fourier transform of $\rho$ and $K_{\textrm{homo}}$  are
\begin{align}
  \hat\rho(\v k_{m_1,m_2,m_3})
  & =
  \frac{V}{M_1M_2M_3}
  \sum_{i_1=0}^{M_1}\sum_{i_2=0}^{M_2}\sum_{i_3=0}^{M_3}
  \rho(\v r_{i_1,i_2,i_3})
  \exp\bigg[
  -2\pi i\bigg(
  \frac{m_1i_1}{M_1} + \frac{m_2i_2}{M_2} + \frac{m_3i_3}{M_3}
  \bigg)
  \bigg],
\end{align}
and
\begin{align}\label{eqn:f-kernel-homo}
  \hat K_{\textrm{homo}}(\v k)
  &=
  \int_{\mathbb R^3}K_{\textrm{homo}}(\v s)e^{-2\pi i\v k\cdot\v s}\,\d d\v s.
\end{align}
Most of the short-range
interactions are isotropic, i.e. $K_{\textrm{homo}}(\v r) =
K_{\textrm{homo}}(r)$, then the three-dimensional integral in
Eqn. \eqref{eqn:f-kernel-homo} can be simplified to a one-dimensional
integral:
\begin{align}\label{eqn:kernel-inte}
  \hat{K}_{\textrm{homo}}(\v k)
  =
  % \int_{\mathbb R^3}K_{\textrm{homo}}(r)e^{-2\pi i\v k\cdot\v r}\d d\v r
  % =
  \int_{r_c}^\infty \frac{2 s}k\, K_{\textrm{homo}}(s) \sin(2\pi ks)\,\d ds,
\end{align}
which can be calculated by numerical
integration from $r_c$ to a large upper bound, saying $r_u$. \recheck{Perfect
convergence is achieved by using $r_u = 30\sigma$
for the dispersion energy $-4\epsilon (\sigma / r)^6$, where $\epsilon$
and $\sigma$ are the energy and length units.}
Since integral~\eqref{eqn:kernel-inte} is independent of the
density of the system, it is calculated only once and stored.

@@

A basic observation in an inhomogeneous system is that the RMS force
error is not uniformly distributed. The difference between the maximum
and minimum error can be more than one order of
magnitude
% in one of the cases we studied,
(see Fig. \ref{fig:tmp2}). Traditionally, a very large uniform cut-off radius is
used to reduce the maximum error in the system to reach convergent
simulation results~\cite{ismail2007application}.
\recheck{
  However, since the computational effort scales as the
  cube of the cut-off radius, a slightly larger  cut-off
  will notably increase the computational effort.
  For example, increasing the cut-off radius
  from $4\sigma$ to $5\sigma$ will cost
  twice the computational effort.}
Therefore, a huge
amount of computational effort is wasted in the low-error region. The
idea of the adaptive cut-off radius simulation is to use larger
cut-off radius for particles in the high-error region to control the
force error, while to use smaller cut-off radius for particles in the
low-error region to save computational cost.  The effectiveness of this
method relies on a deliberated cut-off choice for each particle so
that the force error is uniformly distributed across the simulation
region.
 In order to determine the cut-off radius, we should estimate the RMS
force error at least once for each particle, and the computational
cost of each force error estimate is $\mathcal O(N \log N)$,
so the total computational expense of 
this method is at least $\mathcal O(N^2\log N)$, which is
prohibitive to the on-the-fly application.

@@

To make the adaptive cut-off method feasible, we assume:
\begin{enumerate}
\item The space is uniformly divided in to small bins, the size of which is the
  same as those mentioned in the section \ref{sec:tmp1.3}.  The
  particles falling in the same bin are assigned the same cut-off
  radius. In other words, the cut-off radius is chosen for each bin
  rather than for each particles.
  % The cut-off
  % radius is assumed to be constant within a bin, no matter the
  % position of the partile.
\item The cut-off radius can only be chosen from a discrete series of
  monotonically increasing values between a maximum 
  $r_c^{\textrm{max}}$ and a minimum 
  $r_c^{\textrm{min}}$: $\{\,r_c^{\textrm{min}} = r_c^0 < r_c^1 <
  \cdots < r_c^{M-1} <r_c^M = r_c^{\textrm{max}}\}$.  We call it the
  \emph{candidate set} of the cut-off radius.  For convenience but not
  necessarily, the cut-off radii in the candidate set is uniformly increasing, i.e.
  \begin{align}
    r_c^m = r_c^0 + r_c^{\textrm{step}} m, \quad 0 \leq m \leq M.
  \end{align}
  For example, a possible candidate set is $\{2.5\,\sigma,\
  3.0\,\sigma,\ 3.5\,\sigma,\ 4.0\,\sigma,\ \cdots,\ 10.0\,\sigma\}$,
  with $r_c^{\textrm{max}} = 10.0\sigma$, $r_c^{\textrm{min}} =
  2.5\sigma$ and $r_c^{\textrm{step}} = 0.5 \sigma$.
\end{enumerate}
For each cut-off radius in the candidate set, the force error
$\mathcal E(\v r)$ is calculated by the estimate Eqn.~\eqref{eqn:tmp5}.
Notice the error is calculated by the
fast algorithm described in section~\ref{sec:tmp1.3}, so it is
defined on the bins.  For a predetermined error control level
$\mathcal E_{\textrm{C}}$, a satisfactory cut-off radius for a bin is
chosen such that it is the smallest one satisfying the precision
constraint $\mathcal E(\v r_{i_1,i_2,i_3}) \leq \mathcal
E_{\textrm{C}}$. The computational cost of this algorithm is $\mathcal
O(MN\log N)$.  The idea of this algorithm is based on the observation
that the cut-off radius may not be calculated at a very high
precision.
% a slightly (at most $r_c^{\textrm{step}}$) larger
% cut-off radius
\recheck{
  Because if $r_c^{\textrm{step}}$ is sufficiently small,
  increasing the cut-off radius by no more than $r_c^{\textrm{step}}$
will not waste too much computational cost.
Here we recommend that $r_c^{\textrm{step}}$ should be no more than
0$.25\sigma$ that increases
the computational effort by 27\% at $r_c = 3.0\sigma$.}
Due to the
thermodynamic fluctuation of the particle density, the resulting cut-off
distribution $r_c(\v r_{i_1,i_2,i_3})$ (as a function of bin
position) needs to be refined by the following scheme:
\begin{align}\label{eqn:rc-refine}
  r^{(1)}_c(\v r_{i_1,i_2,i_3}) = \max_{j_\alpha\in I}\,r_c(\v r_{i_1+j_1,i_2+j_2,i_3+j_3}),
  \quad I = \{-1, 0, 1\},
\end{align}
so that the fluctuating large-error regions are safely covered by bins
with large cut-off radius.

@@

The direct simulation of the liquid-vapor equilibrium is a very
typical inhomogeneous system, which is widely used to study the phase
coexistence properties, for example, the equilibrium liquid/vapor
density, and the surface tension. The system studied in the present
paper contains 16,000 identical standard Lennard-Jones particles that are
separated into liquid phase in the center of the periodic box and vapor phase
surrounded (see Fig.~\ref{fig:tmp1}).  The conventional Lennard-Jones
unit system is employed: the unit of length, energy, mass and time are
denoted by $\epsilon$, $\sigma$, $m$ and $\tau$ ($\tau =
\sigma\sqrt{m/\varepsilon}$), respectively.  In the rest of this paper,
all quantities are
rescaled by the Lennard-Jones units and added the superscript ``$\ast$''.
For example, $r^\ast = r / \sigma$, $T^\ast = k_BT / \epsilon$,
$P^\ast = P\cdot \sigma^3 / \epsilon$ and $\mathcal E^\ast = \mathcal
E\cdot\sigma/\varepsilon$.  The MD time step {is} $\Delta t^\ast =
0.005$. The simulations last for $3\times 10^6$ time steps. The first
$1\times 10^6$ steps are discarded. The quantities of interest are
sampled every 100 time steps. The blocking average
method~\cite{flyvbjerg1989error} is applied to estimate the
statistical uncertainty of the autocorrelated data with 95\%
confidence level. The NVT ensemble is generated by the Nos\`e-Hoover
thermostat~\cite{nose1984molecular, hoover1985canonical}.  The size of
the simulation box is $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 21\times 21$, which permits a maximum cut-off radius of
$r_c^\ast = 10$, and eliminates the finite size effect reported in
literatures \cite{chen1995area, orea2005oscillatory,
  biscay2009calculation}.
The system is divided into bins of size $1\times1\times1$ to perform the fast error
estimate method described in section \ref{sec:tmp1.3}.
In all simulations, the maximum cut-off
radius considered is $10$, but the reference results for comparison
are obtained by simulations carried out in a 24,000 particles system
with simulation box $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 27\times 27$, and a uniform cut-off radius of $r_c^\ast =
13$. \recheck {We find that (not shown here)
  the bigger system leads to the same equilibrium
  quantities as the smaller system at the same cut-off radius, so
  it is reasonable to
  use it to produce the better converged data at $r_c^\ast=13$ and compare them with
  those obtained in the smaller system.
}

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{fig/t0.85-n16000-rc07.5uni/confout.eps}
  \caption{The snapshot of the 16000 Lennard-Jones particle system in
    liquid-vapor equilibrium at $T^\ast=0.85$ with a uniform cut-off
    radius of $r_c^\ast = 7.5$.}
  \label{fig:tmp1}
\end{figure}

@@

\begin{figure}
  \centering
  % \includegraphics[]{fig/t0.85-n16000-rc07.5uni/confout-02.eps}
  \includegraphics[]{fig/t0.85-n16000-rc07.5uni/error.uniform.eps}
  \caption{The RMS force error distribution of the 16000 Lennard-Jones
    particle system at $T^\ast=0.85$ with a uniform cut-off radius of
    $r_c^\ast = 7.5$. The green squares present the real error. The
    solid blue line is the error estimate by \eqref{eqn:tmp5}. The
    dashed blue line is the homogeneity error contribution, while the
    dotted blue line is the inhomogeneity error contribution. The
    density distribution of the system is presented by red solid line
    for reference.  All properties are properly averaged on the $y$
    and $z$ direction, therefore, the profiles are plotted along the
    $x$ direction.  }
  \label{fig:tmp2}
\end{figure}

% The Fig. \ref{fig:tmp2} inspires us to develop the adaptive cut-off
% method: in the bulk regions, the error is much smaller, therefore, it
% is a waste to use the same cut-off radius as the interfacial
% regions.
As mentioned before, the key idea of the adaptive cut-off simulation
is to use smaller cut-off
radius in the bulk regions, so that the RMS force error is uniformly
distributed over the simulation region. In Fig. \ref{fig:tmp3}, we
present the adapted cut-off radius by a solid red line and all errors
by the same notations as in the Fig. \ref{fig:tmp2}. The control error
$\mathcal E^\ast_{\textrm{C}} = 0.0045$ is almost the same as the maximum
error in the uniform cut-off simulation with $r_c^\ast=7.5$.
\recheck{
  In the rest of the paper, when we mention that ``the cut-off of
  the adaptive cut-off method is $r_c^\ast$'', we mean the
  control error is the maximum error of a uniform cut-off simulation
  using that cut-off $r_c^\ast$. In all the cases considered by
  the present paper, the maximum of
  the adapted cut-off radius is almost the same as $r_c^\ast$,
  and is sometimes marginally larger (no more than $r_c^\textrm{step}$).
}
The
cut-off distribution is calculated every 20,000 time steps, and is
refined twice by Eqn.~\eqref{eqn:rc-refine}. In the interfacial
regions, the same cut-off radius as the uniform cut-off simulation,
namely $7.5$, is used.  In the bulk
liquid and vapor regions, the cut-off radii are $4.75$ and $3.5$,
which reduce the computational cost
by a factor of 4 and 10, respectively.
The error is uniformly distributed, except in the
liquid region where the real error is somewhat lower.
In contrast with the uniform
cut-off case, i.e. Fig. \ref{fig:tmp2}, the error estimate for the
vapor region is very sharp.
The similar phenomenon is observed in
Ref.~\cite{wang2011}.
The same as the uniform cut-off simulation, the
homogeneity error dominates in the bulk regions, while the
inhomogeneity error dominates in the interfacial region.

\begin{figure}
  \centering
  % \includegraphics[scale=1]{fig/t0.85-n16000-rc07.5uni/confout-02.eps}
  \includegraphics[]{fig/t0.85-n16000-adapt-e0.0045-extend/rcut.and.error.eps}
  \caption{ The RMS force error distribution of the 16000
    Lennard-Jones particle system at $T^\ast=0.85$ with adaptive
    cut-off radius. The control error $\mathcal E^\ast_{\textrm{C}}$
    is the same as the maximum error in the uniform cut-off simulation
    with $r_c^\ast=7.5$.  The resulting cut-off radius distribution is
    presented by the red line. All other notations are the same as
    Fig.~\ref{fig:tmp2}.}
  \label{fig:tmp3}
\end{figure}

@@

\begin{figure}
  \centering
  % \includegraphics[scale=1]{fig/t0.85-n16000-rc07.5uni/confout-02.eps}
  \includegraphics[]{fig/t0.85-n16000-fcorr-rc07.5-feq0200/fcorr.and.error.eps}
  \caption{ The RMS force error distribution of the 16000
    Lennard-Jones particle system at $T^\ast=0.85$ with long-range
    force correction, i.e. eqn. \eqref{eqn:define-fcorr}. The cut-off
    radius is $r^\ast_c = 7.5$. The $x$ component of the correction
    force is presented by the solid red line. The green squares
    present the real error. The solid blue line presents the estimated
    homogeneity error. }
  \label{fig:tmp4}
\end{figure}

>>>>

\begin{abstract}
  {
    In the present paper, we develop an accurate error estimate of the
    non-bonded short-range interaction for the inhomogeneous molecular
    systems.
    The root mean square force error is proved to be decomposed into
    three additive parts: the homogeneity error, the inhomogeneity error
    and the correlation error.
    % The homogeneity error is dominant
    % in the bulk regions.
    The magnitude of the inhomogeneity error, which
    is dominant in the interfacial regions, 
    can be more than one order of magnitude larger than the homogeneity
    error.
    % This is the reason why a direct simulation may waste
    % a huge amount of computational effort.
    This is the reason why a standard simulation with fixed
    cut-off radius is either less accurate if the cut-off is too small,
    or wastes considerable computational effort if the cut-off is too large.
    Therefore, based on the error estimate, the adaptive cut-off
    and long-range force correction methods are proposed to
    boost the efficiency and accuracy of the simulation, respectively.
    The
    effectiveness of the proposed methods are demonstrated by
    molecular dynamics 
    simulations of the liquid-vapor equilibrium and the nanoscale particle
    collision. Different roles of the homogeneity error and inhomogeneity
    error are also discussed.
  }
\end{abstract}

@@

The naive calculation of the convolutions in the error estimate \eqref{eqn:tmp5} requires
a computational cost of $\mathcal O(N^2)$. To estimate the force error
on the fly, the computational cost of the estimate should be reduced to
at least $\mathcal O(N\log N)$.  This is achieved by  using the fast Fourier transform. We
assume the simulation box is uniformly divided in to small bins.  The
positions of these bins are
\begin{align}
  \v r_{i_1, i_2, i_3} & =
  \frac{i_1}{M_1}\v a_1 + 
  \frac{i_2}{M_2}\v a_2 + 
  \frac{i_3}{M_3}\v a_3,\qquad 0\leq i_\alpha< M_\alpha
\end{align}
where $\v a_\alpha$ are box vectors, and $M_\alpha$ is the number of
divisions on each direction. In the reciprocal space, a corresponding
lattice is set up:
\begin{align}
  \v k_{m_1, m_2, m_3} & =
  m_1\v a_1^\ast +
  m_2\v a_2^\ast +
  m_3\v a_3^\ast, \qquad 0\leq m_\alpha<M_\alpha
\end{align}
where $\v a_\alpha^\ast$ are reciprocal box vectors defined by $\v
a_\alpha\cdot\v a_\beta^\ast = \delta_{\alpha\beta}$.  In each bin the
error and the particles number density are assumed to be constant,
then, a Fourier expansion can be obtained:
\begin{align}\nonumber
  \mathcal E^2_{\textrm{homo}}(\v r_{i_1,i_2,i_3})
  & =
  \frac1V\sum_{m_1=-\infty}^\infty\sum_{m_2=-\infty}^\infty\sum_{m_3=-\infty}^\infty
  [{\mathcal E}^2_{\textrm{homo}}]^{\wedge}(\v k_{m_1,m_2,m_3})
  % e^{2\pi i\v k_{m_1,m_2,m_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{m_1i_1}{M_1} + \frac{m_2i_2}{M_2} + \frac{m_3i_3}{M_3}
  \bigg)
  \bigg] \\\label{eqn:tmp13}
  & =
  \frac1V\sum_{m_1=0}^{M_1}\sum_{m_2=0}^{M_2}\sum_{m_3=0}^{M_3}
  \hat{K}_{\textrm{homo}}(\v k'_{m_1,m_2,m_3})\:
  \hat{\rho}(\v k_{m_1,m_2,m_3})
  % e^{2\pi i\v k_{m_1,m_2,m_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{m_1i_1}{M_1} + \frac{m_2i_2}{M_2} + \frac{m_3i_3}{M_3}
  \bigg)
  \bigg].
\end{align}
Here $K_{\textrm{homo}}(\v r)$ denotes $\vert\v f^c(\v r)\vert^2$.
The wedge notation $\wedge$ means the Fourier transform. The prime on $\v k_{m_1,m_2,m_3}$ means
that the periodic image $m_\alpha - M_\alpha$ should use instead of
$m_\alpha$ when $m_\alpha \geq M_\alpha/2$.
The second equation in \eqref{eqn:tmp13}
holds due to the fact:
$[\mathcal E^2]^\wedge = [K_{\textrm{homo}}\ast\rho]^\wedge = \hat{K}_{\textrm{homo}} \times \hat\rho$. The Fourier transform of $\rho$ and $K_{\textrm{homo}}$  are
\begin{align}
  \hat\rho(\v k_{m_1,m_2,m_3})
  & =
  \frac{V}{M_1M_2M_3}
  \sum_{i_1=0}^{M_1}\sum_{i_2=0}^{M_2}\sum_{i_3=0}^{M_3}
  \rho(\v r_{i_1,i_2,i_3})
  \exp\bigg[
  -2\pi i\bigg(
  \frac{m_1i_1}{M_1} + \frac{m_2i_2}{M_2} + \frac{m_3i_3}{M_3}
  \bigg)
  \bigg],
\end{align}
and
\begin{align}\label{eqn:f-kernel-homo}
  \hat K_{\textrm{homo}}(\v k)
  &=
  \int_{\mathbb R^3}K_{\textrm{homo}}(\v s)e^{-2\pi i\v k\cdot\v s}\,\d d\v s.
\end{align}
Most of the short-range
interactions are isotropic, i.e. $K_{\textrm{homo}}(\v r) =
K_{\textrm{homo}}(r)$, then the three-dimensional integral in
Eqn. \eqref{eqn:f-kernel-homo} can be simplified to a one-dimensional
integral:
\begin{align}\label{eqn:kernel-inte}
  \hat{K}_{\textrm{homo}}(\v k)
  =
  % \int_{\mathbb R^3}K_{\textrm{homo}}(r)e^{-2\pi i\v k\cdot\v r}\d d\v r
  % =
  \int_{r_c}^\infty \frac{2 s}k\, K_{\textrm{homo}}(s) \sin(2\pi ks)\,\d ds,
\end{align}
which can be calculated by numerical
integration from $r_c$ to a large upper bound, saying $r_u$. {Perfect
convergence is achieved by using $r_u = 30\sigma$
for the dispersion energy $-4\epsilon (\sigma / r)^6$, where $\epsilon$
and $\sigma$ are the energy and length units.}
Since integral~\eqref{eqn:kernel-inte} is independent of the
density of the system, it is calculated only once and stored.

@@

A basic observation in an inhomogeneous system is that the RMS force
error is not uniformly distributed. The difference between the maximum
and minimum error can be more than one order of
magnitude
% in one of the cases we studied,
(see Fig. \ref{fig:tmp2}). Traditionally, a very large uniform cut-off radius is
used to reduce the maximum error in the system to reach convergent
simulation results~\cite{ismail2007application}.
{
  However, since the computational effort scales as the
  cube of the cut-off radius, a slightly larger  cut-off
  will notably increase the computational effort.
  For example, increasing the cut-off radius
  from $4\sigma$ to $5\sigma$ will cost
  twice the computational effort.}
Therefore, a huge
amount of computational effort is wasted in the low-error region. The
idea of the adaptive cut-off radius simulation is to use larger
cut-off radius for particles in the high-error region to control the
force error, while to use smaller cut-off radius for particles in the
low-error region to save computational cost.  The effectiveness of this
method relies on a deliberated cut-off choice for each particle so
that the force error is uniformly distributed across the simulation
region.
 In order to determine the cut-off radius, we should estimate the RMS
force error at least once for each particle, and the computational
cost of each force error estimate is $\mathcal O(N \log N)$,
so the total computational expense of 
this method is at least $\mathcal O(N^2\log N)$, which is
prohibitive to the on-the-fly application.

@@

To make the adaptive cut-off method feasible, we assume:
\begin{enumerate}
\item The space is uniformly divided in to small bins, the size of which is the
  same as those mentioned in the section \ref{sec:tmp1.3}.  The
  particles falling in the same bin are assigned the same cut-off
  radius. In other words, the cut-off radius is chosen for each bin
  rather than for each particles.
  % The cut-off
  % radius is assumed to be constant within a bin, no matter the
  % position of the partile.
\item The cut-off radius can only be chosen from a discrete series of
  monotonically increasing values between a maximum 
  $r_c^{\textrm{max}}$ and a minimum 
  $r_c^{\textrm{min}}$: $\{\,r_c^{\textrm{min}} = r_c^0 < r_c^1 <
  \cdots < r_c^{M-1} <r_c^M = r_c^{\textrm{max}}\}$.  We call it the
  \emph{candidate set} of the cut-off radius.  For convenience but not
  necessarily, the cut-off radii in the candidate set is uniformly increasing, i.e.
  \begin{align}
    r_c^m = r_c^0 + r_c^{\textrm{step}} m, \quad 0 \leq m \leq M.
  \end{align}
  For example, a possible candidate set is $\{2.5\,\sigma,\
  3.0\,\sigma,\ 3.5\,\sigma,\ 4.0\,\sigma,\ \cdots,\ 10.0\,\sigma\}$,
  with $r_c^{\textrm{max}} = 10.0\sigma$, $r_c^{\textrm{min}} =
  2.5\sigma$ and $r_c^{\textrm{step}} = 0.5 \sigma$.
\end{enumerate}
For each cut-off radius in the candidate set, the force error
$\mathcal E(\v r)$ is calculated by the estimate Eqn.~\eqref{eqn:tmp5}.
Notice the error is calculated by the
fast algorithm described in section~\ref{sec:tmp1.3}, so it is
defined on the bins.  For a predetermined error control level
$\mathcal E_{\textrm{C}}$, a satisfactory cut-off radius for a bin is
chosen such that it is the smallest one satisfying the precision
constraint $\mathcal E(\v r_{i_1,i_2,i_3}) \leq \mathcal
E_{\textrm{C}}$. The computational cost of this algorithm is $\mathcal
O(MN\log N)$.  The idea of this algorithm is based on the observation
that the cut-off radius may not be calculated at a very high
precision.
% a slightly (at most $r_c^{\textrm{step}}$) larger
% cut-off radius
{
  Because if $r_c^{\textrm{step}}$ is sufficiently small,
  increasing the cut-off radius by no more than $r_c^{\textrm{step}}$
will not waste too much computational cost.
Here we recommend that $r_c^{\textrm{step}}$ should be no more than
0$.25\sigma$ that increases
the computational effort by 27\% at $r_c = 3.0\sigma$.}
Due to the
thermodynamic fluctuation of the particle density, the resulting cut-off
distribution $r_c(\v r_{i_1,i_2,i_3})$ (as a function of bin
position) needs to be refined by the following scheme:
\begin{align}\label{eqn:rc-refine}
  r^{(1)}_c(\v r_{i_1,i_2,i_3}) = \max_{j_\alpha\in I}\,r_c(\v r_{i_1+j_1,i_2+j_2,i_3+j_3}),
  \quad I = \{-1, 0, 1\},
\end{align}
so that the fluctuating large-error regions are safely covered by bins
with large cut-off radius.

@@

The direct simulation of the liquid-vapor equilibrium is a very
typical inhomogeneous system, which is widely used to study the phase
coexistence properties, for example, the equilibrium liquid/vapor
density, and the surface tension. The system studied in the present
paper contains 16,000 identical standard Lennard-Jones particles that are
separated into liquid phase in the center of the periodic box and vapor phase
surrounded (see Fig.~\ref{fig:tmp1}).  The conventional Lennard-Jones
unit system is employed: the unit of length, energy, mass and time are
denoted by $\epsilon$, $\sigma$, $m$ and $\tau$ ($\tau =
\sigma\sqrt{m/\varepsilon}$), respectively.  In the rest of this paper,
all quantities are
rescaled by the Lennard-Jones units and added the superscript ``$\ast$''.
For example, $r^\ast = r / \sigma$, $T^\ast = k_BT / \epsilon$,
$P^\ast = P\cdot \sigma^3 / \epsilon$ and $\mathcal E^\ast = \mathcal
E\cdot\sigma/\varepsilon$.  The MD time step {is} $\Delta t^\ast =
0.005$. The simulations last for $3\times 10^6$ time steps. The first
$1\times 10^6$ steps are discarded. The quantities of interest are
sampled every 100 time steps. The blocking average
method~\cite{flyvbjerg1989error} is applied to estimate the
statistical uncertainty of the autocorrelated data with 95\%
confidence level. The NVT ensemble is generated by the Nos\`e-Hoover
thermostat~\cite{nose1984molecular, hoover1985canonical}.  The size of
the simulation box is $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 21\times 21$, which permits a maximum cut-off radius of
$r_c^\ast = 10$, and eliminates the finite size effect reported in
literatures \cite{chen1995area, orea2005oscillatory,
  biscay2009calculation}.
The system is divided into bins of size $1\times1\times1$ to perform the fast error
estimate method described in section \ref{sec:tmp1.3}.
In all simulations, the maximum cut-off
radius considered is $10$, but the reference results for comparison
are obtained by simulations carried out in a 24,000 particles system
with simulation box $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 27\times 27$, and a uniform cut-off radius of $r_c^\ast =
13$.  {We find that (not shown here)
  the bigger system leads to the same equilibrium
  quantities as the smaller system at the same cut-off radius, so
  it is reasonable to
  use it to produce the better converged data at $r_c^\ast=13$ and compare them with
  those obtained in the smaller system.
}

\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth]{figs/confout.eps}
  \caption{The snapshot of the 16000 Lennard-Jones particle system in
    liquid-vapor equilibrium at $T^\ast=0.85$ with a uniform cut-off
    radius of $r_c^\ast = 7.5$.}
  \label{fig:tmp1}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[]{figs/error.uniform.eps}
  \caption{The RMS force error distribution of the 16000 Lennard-Jones
    particle system at $T^\ast=0.85$ with a uniform cut-off radius of
    $r_c^\ast = 7.5$. The green squares present the real error. The
    solid blue line is the error estimate by \eqref{eqn:tmp5}. The
    dashed blue line is the homogeneity error contribution, while the
    dotted blue line is the inhomogeneity error contribution. The
    density distribution of the system is presented by red solid line
    for reference.  All properties are properly averaged on the $y$
    and $z$ direction, therefore, the profiles are plotted along the
    $x$ direction.  }
  \label{fig:tmp2}
\end{figure}

% The Fig. \ref{fig:tmp2} inspires us to develop the adaptive cut-off
% method: in the bulk regions, the error is much smaller, therefore, it
% is a waste to use the same cut-off radius as the interfacial
% regions.
As mentioned before, the key idea of the adaptive cut-off simulation
is to use smaller cut-off
radius in the bulk regions, so that the RMS force error is uniformly
distributed over the simulation region. In Fig. \ref{fig:tmp3}, we
present the adapted cut-off radius by a solid red line and all errors
by the same notations as in the Fig. \ref{fig:tmp2}. The control error
$\mathcal E^\ast_{\textrm{C}} = 0.0045$ is almost the same as the maximum
error in the uniform cut-off simulation with $r_c^\ast=7.5$.
{
  In the rest of the paper, when we mention that ``the cut-off of
  the adaptive cut-off method is $r_c^\ast$'', we mean the
  control error is the maximum error of a uniform cut-off simulation
  using that cut-off $r_c^\ast$. In all the cases considered by
  the present paper, the maximum of
  the adapted cut-off radius is almost the same as $r_c^\ast$,
  and is sometimes marginally larger (no more than $r_c^\textrm{step}$).
}
The
cut-off distribution is calculated every 20,000 time steps, and is
refined twice by Eqn.~\eqref{eqn:rc-refine}. In the interfacial
regions, the same cut-off radius as the uniform cut-off simulation,
namely $7.5$, is used.  In the bulk
liquid and vapor regions, the cut-off radii are $4.75$ and $3.5$,
which reduce the computational cost
by a factor of 4 and 10, respectively.
The error is uniformly distributed, except in the
liquid region where the real error is somewhat lower.
In contrast with the uniform
cut-off case, i.e. Fig. \ref{fig:tmp2}, the error estimate for the
vapor region is very sharp.
The similar phenomenon is observed in
Ref.~\cite{wang2011}.
The same as the uniform cut-off simulation, the
homogeneity error dominates in the bulk regions, while the
inhomogeneity error dominates in the interfacial region.

\begin{figure}
  \centering
  \includegraphics[]{figs/rcut.and.error.eps}
  \caption{ The RMS force error distribution of the 16000
    Lennard-Jones particle system at $T^\ast=0.85$ with adaptive
    cut-off radius. The control error $\mathcal E^\ast_{\textrm{C}}$
    is the same as the maximum error in the uniform cut-off simulation
    with $r_c^\ast=7.5$.  The resulting cut-off radius distribution is
    presented by the red line. All other notations are the same as
    Fig.~\ref{fig:tmp2}.}
  \label{fig:tmp3}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[]{figs/fcorr.and.error.eps}
  \caption{ The RMS force error distribution of the 16000
    Lennard-Jones particle system at $T^\ast=0.85$ with long-range
    force correction, i.e. eqn. \eqref{eqn:define-fcorr}. The cut-off
    radius is $r^\ast_c = 7.5$. The $x$ component of the correction
    force is presented by the solid red line. The green squares
    present the real error. The solid blue line presents the estimated
    homogeneity error. }
  \label{fig:tmp4}
\end{figure}
